---
title: "Predictive modeling exam P2"
author: "Reece Wooten"
date: "8/1/2017"
output: html_document
---

```{r}
data=read.csv("MidCity.csv")
set.seed(1)

```
## Problem 2 From the Exam
#### 1. 
##### Is there a premium for brick houses everything else being equal?
```{r}
data=read.csv("MidCity.csv")
data$Nbhd=factor(data$Nbhd)
fit=lm(Price~.,data=data)
summary(fit)
set.seed(1)


```
* Holding all else equal, the variable for Brick houses is statistically significant and has a positive value. Which indicates that on average and all else equal, brick houses increase price by $17,313.540 compared to non brick houses. 

#### 2. 
##### Is there a premium for houses in neighborhood 3?
* After converting the neighborhood variable into factor levels, on average and all else equal, neighborhood 3 is $20,534 more expensive than neighborhood 1. 


#### 3. 
##### Is there an extra premium for brick houses in neighborhood 3?
```{r}
set.seed(1)

fit=lm(Price~Home+Nbhd+Offers+SqFt+Brick+Bedrooms+Bathrooms+(Nbhd*Brick),data=data)
summary(fit)
rm(list=ls())
```
* The interaction between neighborhood 3 and the brick variable allows the model to check for the premium in brick houses in neighborhood 3 compared to neighborhood 1. The variable is statistically significant at the 5% level and on average holding all else equal a brick house in neighborhood 3 will cost 12,019 more than a brick house in neighborhood 1. 

#### 4. 
##### For the purposes of prediction could you combine the neighborhoods 1 and 2 into a single “older” neighborhood?
* From the previous questions output, we see that neighborhood two does not have a 5% statistically significant effect on the price of a house compared to neighborhood one. Because of this result, we could combine neighborhood one and two since we can assume the effect of neighborhood two on price is no different than neighborhood one. 
